---
title: "Doing parentage on black-and-yellow rockfish"
output: html_notebook
 toc: true
---

This is a first stab at parentage in black-and-yellow rockfish, based on the analysis Eric did for kelps.  In a previous document, `./Rmd/01-aggregate-genos-and-gsi-to-species.Rmd` 
I aggregated genotypes and filtered high missers and did GSI to identify species amongst the juveniles
and the adults. I saved those results and use them here...

To start off with, let's load data and libs:
```{r load-stuff}
library(tidyverse)
library(CKMRsim)
library(stringr)


meta <- readRDS("../data/processed/meta-data-tibble.rds")
genos <- readRDS("../extdata/processed/genos-aggregated-and-no-hi-missers.rds")
juvie_species <- readRDS("../data/processed/juvie-gsi-sim-assignments.rds")   # gsi assignments of juvenile fish
adult_species <- readRDS("../data/processed/adult-gsi-sim-self-assignments.rds")   # gsi self-assignments of adult fish
```

## Selecting our fish and designating allele nomenclature

We gotta pick out the _S. chrysomelas_.  For adults that is mostly done, but there are some
obvious misassigmnments, so we use the GSI results for both adults and juvies.  
```{r define-black-yellows}
blackyellow_ad <- adult_species %>% 
  filter(species == "chrysomelas",
         score > 99) %>%
  select(NMFS_DNA_ID) %>% unlist() %>% unname()

blackyellow_juv <- juvie_species %>%
  filter(species == "chrysomelas",
         score > 99) %>%
  select(NMFS_DNA_ID) %>% unlist() %>% unname

```

That is `r length(blackyellow_ad)` adults and `r length(blackyellow_juv)` juveniles.  
With those we filter down the genos to the ones that we want, and then we
can get it into the format required for CKMR.
```{r blackyellow-genos}
blackyellow_genos <- genos %>%
  filter(NMFS_DNA_ID %in% c(blackyellow_ad, blackyellow_juv))

# we will use this some more
byg2 <- blackyellow_genos %>% 
  select(NMFS_DNA_ID, locus, allele) %>%
  mutate(Chrom = "GTseq") %>% 
  mutate(Pos = as.integer(factor(locus, levels = unique(locus)))) %>%
  rename(Locus = locus,
         Allele = allele) %>%
  select(NMFS_DNA_ID, Chrom, Locus, Pos, Allele) %>%
  ungroup()

# get the allele freqs
byg_ckmr_markers <- byg2 %>%
  filter(!is.na(Allele)) %>% # it is vital to filter out the NAs at this stage
  group_by(Chrom, Locus, Pos, Allele) %>%
  summarise(counts = n()) %>%
  group_by(Locus, Pos) %>%
  mutate(Freq = counts / sum(counts)) %>%
  select(-counts) %>%
  mutate(AlleIdx = 1,
         LocIdx = 1) %>%
  reindex_markers(.)
```
We have a total `r nrow(byg_ckmr_markers)` alleles for black-and-yellow rockfish in this data set.

## Doing CKMR simulations to assess power

Really simply let's just do this to see what our power for parentage ought to look like.
```{r CKMR-create}
CK_by <- create_ckmr(byg_ckmr_markers, kappa_matrix = kappas[c("PO", "FS", "HS", "U"), ])

# see what it tells us when we print it:
CK_by
```

Then do the simulations to see what the false positive rates are for finding parents when the truth
is U or FS:
```{r CKMR-sim-and-sample}
Qs <- simulate_Qij(C = CK_by, froms = c("PO", "FS", "HS", "U"), tos = c("PO", "FS", "HS", "U"), reps = 10^4)

# then do the importance sampling to get the FPRs
mc_sample_simple(Qs, nu = "PO", de = c("U", "FS"), tr = c("U", "FS"), method = "both")
```

We can come back to that later to see about finding Full Sibs.

## Making genotype matrices
```{r}
blackyellow_haps <- blackyellow_genos %>%
  filter(!is.na(allele)) %>%  # once again, it is critical to remove these at this point
  select(NMFS_DNA_ID, locus, gene_copy, allele) %>%
  rename(Locus = locus, Allele = allele)

blackyellow_idx_frame <- byg_ckmr_markers %>%
  select(Locus, Allele, LocIdx, AlleIdx) %>%
  group_by(Locus) %>%
  mutate(NumA = n()) %>%  # get the number of alleles at each locus
  ungroup() %>%
  left_join(blackyellow_haps, .)  %>%  # join the alle_idx's onto the actual genotype data
  select(NMFS_DNA_ID, Locus, gene_copy, LocIdx, NumA, AlleIdx) %>%
  spread(key = gene_copy, value = AlleIdx) %>%
  mutate(GenoIdx = index_ab(a = `1`, b = `2`, A = NumA))

# make a matrix of genotype integers 
wide_blackyellow <- blackyellow_idx_frame %>%
  select(NMFS_DNA_ID, LocIdx, GenoIdx) %>%
  spread(data = ., key = LocIdx, value = GenoIdx)
```

Don't forget to set NA's to 0, and then decrease each value by 1:
```{r make-mat}
rocky_mat <- as.matrix(wide_blackyellow[, -1])
rownames(rocky_mat) <- wide_blackyellow$NMFS_DNA_ID
rocky_mat[is.na(rocky_mat)] <- 0
rocky_mat <- rocky_mat - 1
storage.mode(rocky_mat) <-  "integer"
```

## Looking for duplicated samples

We can quickly look through rocky_mat for pairs of indivs with lots of matching genotypes.
```{r check-for-dupes}
matchers <- pairwise_geno_id(S = rocky_mat, max_miss = 12) %>%
  arrange(num_mismatch) %>%
  mutate(NMFS_DNA_ID_1 = rownames(rocky_mat)[ind1],
         NMFS_DNA_ID_2 = rownames(rocky_mat)[ind2])
matchers
```
We need to deal with these somehow.  But before we do that, I want to use these numbers to come up with
a per-genotype discordance rate.  That should be pretty straightforward:
```{r compute-discordance-rate}
sum(matchers$num_mismatch) / sum(matchers$num_loc)

```
So, to repeat, the discordance rate across all these multiply-genotyped individuals is
`r sum(matchers$num_mismatch) / sum(matchers$num_loc)`---roughly 8 in 1000. Higher than for kelp, but hopefully reasonable. 

### Investigate the read depths of these duplicate mismatches

```{r depths-of-dupies}
# just gonna write this to a file and do it elsewhere
saveRDS(matchers, file = "rds_outputs/blackyellow_matchers.rds")
```

Now, to deal with these duplicates, here is what I will do: I am going to say that anyone with 3 or fewer mismatches is certainly 
the same individual, and we are going to take just one from amongst those pairs/groups.  We use a hasty method:
we toss out all the individuals that appear in both ind1 and ind2, and then all the remaining ind2's.  This 
might throw out one more than we need to, but it is what we do for now.
```{r toss-matchers}
def_same <- matchers %>%
  filter(num_mismatch <= 3)

toss_these <- c(intersect(def_same$ind1, def_same$ind2), def_same$ind2) %>%
  unique()

rocky_mat_tossed <- rocky_mat[-toss_these,]
```

Now we have to break that into two matrices: one for the adults and one for the kids.
This should not be too hard.  We have to deal with the fact that not all of the adults in the meta data
had few enough missing genotypes to be included.
```{r break-stage}
blackyellow_ad_kept <- blackyellow_ad[blackyellow_ad %in% rownames(rocky_mat_tossed)]
blackyellow_juv_kept <- blackyellow_juv[blackyellow_juv %in% rownames(rocky_mat_tossed)]

adult_mat <- rocky_mat_tossed[blackyellow_ad_kept, ]
juvie_mat <- rocky_mat_tossed[blackyellow_juv_kept, ]
```


## Save all those black-and-yellow genotypes for quick analysis of other things
```{r dump-genos}
blackyellow_haps %>%
  filter(NMFS_DNA_ID %in% rownames(rocky_mat_tossed)) %>%
  saveRDS("rds_outputs/blackyellow_genos_used_long.rds", compress = "xz")
```

## Before doing the comparisons, compute internal heterozygosities

We want to have these to make sure we don't have contaminated individauls.
```{r compute-int-hzs}
int_hz <- byg2 %>% 
  group_by(NMFS_DNA_ID, Locus) %>% 
  summarise(hz = Allele[1]!=Allele[2]) %>% 
  filter(!is.na(hz)) %>% 
  group_by(NMFS_DNA_ID) %>% 
  summarise(num_loc = n(), num_hz = sum(hz), fract_hz = num_hz / num_loc)
```
And now let us plot that .
```{r plot-int-hz, fig.width=12}
ggplot(int_hz, aes(x = fract_hz)) +
  geom_density(fill = "orange", alpha = 0.3)
```

So, we would be suspicious of individuals with int_hz > .6 or less than 0.28.

## Doing the comparisons

We need to flatten the CKMR object appropriately
```{r flatten}
# flatten those out
po_flat <- flatten_ckmr(CK_by, "PO")
unrel_flat <- flatten_ckmr(CK_by, "U")

# then compute the log-likelihoods for the parent offspring relationship
po_logl_flat <- po_flat
po_logl_flat$probs <- log(po_flat$probs / unrel_flat$probs)

```

Now we are ready to do these!
```{r do-comps}
idx <- 1:nrow(juvie_mat)
names(idx) <- idx
comps <- lapply(idx, function(i) {
    tmp <- comp_ind_pairwise(S = adult_mat, T = juvie_mat, t = i, values = po_logl_flat$probs, nGenos = po_logl_flat$nGenos, Starts = po_logl_flat$base0_locus_starts)
    #tmp[rev(top_index(tmp$value, 5)), ]  # just take the top 5 from each
    }) %>%
  dplyr::bind_rows(.id = "offspring") %>%
  dplyr::tbl_df() 

# make int_hz something we can join on there
tmp1 <- int_hz %>% rename(id1 = NMFS_DNA_ID, fract_hz1 = fract_hz) %>% select(id1, fract_hz1)
tmp2 <- int_hz %>% rename(id2 = NMFS_DNA_ID, fract_hz2 = fract_hz) %>% select(id2, fract_hz2)

# have a look at that
result <- arrange(comps, desc(value)) %>%
  mutate(number = 1:nrow(.)) %>%
  mutate(id1 = rownames(juvie_mat)[as.integer(offspring)],
         id2 = rownames(adult_mat)[ind]) %>%
  left_join(tmp1) %>%
  left_join(tmp2)

result %>% filter(number <= 1000)

```
None of these values suggest any confident parent-offspring pairs. That's not terribly surprising given the overall small number of comparisons for black-and-yellow adults/juvs.

Try plotting all those pairs, and compare them to the simulated values
```{r plot-all-pairs-density, fig.width = 6}
simlogls <- extract_logls(Q = Qs, numer = c(PO = 1), denom = c(U = 1))
ggplot(result, aes(x = value)) +
  geom_density(colour = "red") +
  geom_density(data = simlogls, aes(x = logl_ratio, fill = true_relat), alpha = 0.25) +
  scale_x_continuous(breaks = seq(-170, 60, by = 20)) +
  xlab("Parent-offspring versus Unrelated Log-Likelihood Ratio")


# save it too
ggsave(filename = "blackyellow_po_logls.pdf", width = 8, height = 4)

```

Put the to 1000 guys on the plot:
```{r distro-plot}
top1000 <- result %>% filter(number <= 1000) %>%
  mutate(yline = -0.001)

ggplot(result, aes(x = value)) +
  geom_density(colour = "red") +
  geom_density(data = simlogls, aes(x = logl_ratio, fill = true_relat), alpha = 0.25) +
  scale_x_continuous(breaks = seq(-170, 60, by = 20)) +
  xlab("Parent-offspring versus Unrelated Log-Likelihood Ratio") +
  geom_jitter(data = top1000, aes(x = value, y = yline), colour = "black", width = 0.0, height = .0003)


ggsave(filename = "black_yellow_po_logls_points.pdf", width = 8, height = 4)
```

The red line for the observed distribution falls above the unrelated distribution, which looks clean.